Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2021-09-26 methods_paper/fabl-paper.tex
update.

\section{Related work}
\label{sec:related}
In this section, we review work that is most related to our proposed work. For surveys of the record linkage literature, see \citep{christen_2012, Herzog_2007, liseo_2013, Winkler2014, Christophides2019, Jurek2019}.

\paragraph{Fellegi and Sunter Method}
Most record linkage techniques are derived from the seminal \citep{fs} paper "A Theory for Record Linkage". The defining
characteristic of their approach was to transform the sets of records,
which often contain text data that is difficult to model, into sets of
comparison vectors governed by parameters that can be more easily
estimated. Concretely, if files \(A\) and \(B\) have \(n_A\) and \(n_B\)
records respectively, and if the files share \(F\) fields in common upon
which to base the linkage, the Fellegi and Sunter approach generates an
\(n_A n_B \times F\) matrix \(\Gamma\), which contains similarity scores
between each pair of records across datasets. We say \(\gamma_{ij}\) is
the comparison vector for record \(i \in A\) and record \(j \in B\),
with \(\gamma_{ij}^f \in \{1, \ldots, L_f\}\) providing their similarity
score on the \(f^{th}\) field. For ease of modeling and computation, we
restrict these similarity scores to be discrete, ordinal variables, and
the construction of these is left to the modeler. We adopt the
convention that \(\gamma_{ij}^f = 1\) corresponds to the highest level
of agreement and \(\gamma_{ij}^f = L_f\) corresponds to the lowest. It
is common to use use binary 1-2 variables to indicate exact matching,
and 1-2-3 variables to provide an option for partial matching. For text
data, we calculate similarity based on Levenstein distance or some other
text similarity score, and bin these scores to integers for use in the
model. \citep{jaro1989} provided guidelines for appropriate binning for the
comparison vectors, but details of implementation are generally context
specific.

The likelihood in the original Fellegi Sunter model is a mixture model
through which each record pair is independently classified as a match or
nonmatch. This independent classification often leads to sets of matches
that break expectations of transitivity, which is undesirable. Although
\citep{jaro1995} devised an procedure to reduce a set of conflicting matches to the
mostly likely set of one-to-one matchings, the abundance of false
matches can bias the estimation of other parameters of the model, and ultimately leads to poor linkage performance.

\citep{sadinle_bayesian_2017} uses the same comparison vector approach as Fellegi and Sunter, but proposed a prior distribution for the set of matches that strictly enforced one-to-one matching without the need for any post-hoc processing procedures. Specifically, in each iteration of his Gibbs sampler, he considers each record \(j\in B\), removes from consideration the records \(i\in A\) that have already been matched, and then samples a potential link. This leads to a sampler that is significantly more accurate than the standard Fellegi-Sunter approach. However,
accounting for these dependencies throughout the linkage process is
computationally burdensome, leaving \texttt{BRL} only suitable for small
to moderate linkage problems.

@@

\section{A scalable model for Bayesian Record Linkage}
\label{sec:model}

In this section, we propose a statistical model suitable for efficient computation in large linkage tasks. Denote two files as \(A\) and \(B\), with \(n_A\) and \(n_B\) records
respectively, and with records indexed as \(i \in \{1, \ldots, n_A\}\)
in \(A\) and \(j \in \{1, \ldots, n_B\}\) in \(B\). Without loss of
generality, label the files such that \(n_A \geq n_B\). We assume there
are no duplicates within files, only across. For each record pair under
consideration, we generate a comparison vector
\(\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}\),
where \(F\) is the number of fields used in the linkage and each
\(\gamma_{ij}^f\) takes on a value \(l \in \{1, \ldots, L_f\}\)
indicating the level agreement between the two records on a specified
field.

@@

>>>>

%\section{Related work}
%\label{sec:related}
%In this section, we review work that is most related to our proposed work. For surveys of the record linkage literature, see \citep{christen_2012, Herzog_2007, liseo_2013, Winkler2014, Christophides2019, Jurek2019}.
%
%\paragraph{Fellegi and Sunter Method}
%Most record linkage techniques are derived from the seminal \citep{fs} paper "A Theory for Record Linkage". The defining
%characteristic of their approach was to transform the sets of records,
%which often contain text data that is difficult to model, into sets of
%comparison vectors governed by parameters that can be more easily
%estimated. Concretely, if files \(A\) and \(B\) have \(n_A\) and \(n_B\)
%records respectively, and if the files share \(F\) fields in common upon
%which to base the linkage, the Fellegi and Sunter approach generates an
%\(n_A n_B \times F\) matrix \(\Gamma\), which contains similarity scores
%between each pair of records across datasets. We say \(\gamma_{ij}\) is
%the comparison vector for record \(i \in A\) and record \(j \in B\),
%with \(\gamma_{ij}^f \in \{1, \ldots, L_f\}\) providing their similarity
%score on the \(f^{th}\) field. For ease of modeling and computation, we
%restrict these similarity scores to be discrete, ordinal variables, and
%the construction of these is left to the modeler. We adopt the
%convention that \(\gamma_{ij}^f = 1\) corresponds to the highest level
%of agreement and \(\gamma_{ij}^f = L_f\) corresponds to the lowest. It
%is common to use use binary 1-2 variables to indicate exact matching,
%and 1-2-3 variables to provide an option for partial matching. For text
%data, we calculate similarity based on Levenstein distance or some other
%text similarity score, and bin these scores to integers for use in the
%model. \citep{jaro1989} provided guidelines for appropriate binning for the
%comparison vectors, but details of implementation are generally context
%specific.
%
%The likelihood in the original Fellegi Sunter model is a mixture model
%through which each record pair is independently classified as a match or
%nonmatch. This independent classification often leads to sets of matches
%that break expectations of transitivity, which is undesirable. Although
%\citep{jaro1995} devised an procedure to reduce a set of conflicting matches to the
%mostly likely set of one-to-one matchings, the abundance of false
%matches can bias the estimation of other parameters of the model, and ultimately leads to poor linkage performance.

@@

\section{Proposed Methodology}
\label{sec:model}

\textcolor{red}{This section needs to be much more clear regarding what is our own contribution and how it differs from the rest of the literature. Right now it's confusing to the reader what comes from Sadinle, FS, etc and why these choices are being made.}

%\citep{sadinle_bayesian_2017} uses the same comparison vector approach as Fellegi and Sunter, but proposed a prior distribution for the set of matches that strictly enforced one-to-one matching without the need for any post-hoc processing procedures. Specifically, in each iteration of his Gibbs sampler, he considers each record \(j\in B\), removes from consideration the records \(i\in A\) that have already been matched, and then samples a potential link. This leads to a sampler that is significantly more accurate than the standard Fellegi-Sunter approach. However,
%accounting for these dependencies throughout the linkage process is
%computationally burdensome, leaving \texttt{BRL} only suitable for small
%to moderate linkage problems.

In this section, we propose the fast beta linkage method for bipartite record linkage, extending the work of \cite{sadinle_bayesian_2017}. Specifically, the author used the same comparison vector approach as Fellegi and Sunter, but proposed a prior distribution for the set of matches that strictly enforced one-to-one matching without the need for any post-hoc processing procedures. In each iteration of his Gibbs sampler, the author considered each record \(j\in \bm{X}_2\), and then removed from consideration the records \(i\in \bm{X}_1\) that have already been matched. Then a potential link is sampled. While this leads to a sampler that is more accurate than the classical Fellegi-Sunter approach, accounting for such dependencies throughout the linkage process is computationally burdensome, and this method is only suitable for small databases.

As such, we propose a framework for efficient computation in large linkage tasks. 
\textcolor{red}{I think my notation is OK from earlier.}

%Denote two files as \(A\) and \(B\), with \(n_A\) and \(n_B\) records
%respectively, and with records indexed as \(i \in \{1, \ldots, n_A\}\)
%in \(A\) and \(j \in \{1, \ldots, n_B\}\) in \(B\). Without loss of
%generality, label the files such that \(n_A \geq n_B\). We assume there
%are no duplicates within files, only across. For each record pair under
%consideration, we generate a comparison vector
%\(\boldsymbol{\gamma}_{ij} = \{\gamma_{ij}^1, \ldots, \gamma_{ij}^F\}\),
%where \(F\) is the number of fields used in the linkage and each
%\(\gamma_{ij}^f\) takes on a value \(l \in \{1, \ldots, L_f\}\)
%indicating the level agreement between the two records on a specified
%field.

@@

\textcolor{red}{Brian: This does not match what Sadinle has, so this needs to be justified regarding the change or if this is a typo. Could you please check this?}
